### 2022-11-22 latex2/grossmanSPAC_homework.tex
Paper 2 update

\documentclass{article}
\usepackage{graphicx}
\usepackage{fullpage}

\parskip 1ex
\parindent 0pt

% an environment for problems that numbers them
\newcounter{exnum}[section]
\newenvironment{problem}{{\vskip 0.1in
   \noindent \bf Problem\addtocounter{exnum}{1}~\arabic{exnum}.~}}{\vskip 0.1in}

% a blunt instrument for making subproblems be (a), (b), ...
\renewcommand{\labelenumi}{(\alph{enumi})}

\begin{document}

\ 
\vfil
\begin{center}
{\Large Some Homework Problems For the Material in ``A Sophomoric
  Introduction to Shared-Memory Parallelism and Concurrency''}
\vfil
\ 
\vfil
For more information, see
http://www.cs.washington.edu/homes/djg/teachingMaterials.
\end{center}
\vfil

\newpage

\begin{problem}{\bf Fork-Join Parallelism: Longest Series}

Consider the problem of finding the longest sequence of some number in
an array of numbers:\\ {\tt longest\_sequence(i,arr)} returns the
longest number of consecutive {\tt i} in {\tt arr}.  For example, if\\
{\tt arr} is {\tt \{2,17,17,8,17,17,17,0,17,1\}} then
{\tt longest\_sequence(17,arr)} is {\tt 3} and 
{\tt longest\_sequence(9,arr)} is {\tt 0}.

\begin{enumerate}
\item In pseudocode, give a parallel fork-join algorithm for
  implementing {\tt longest\_sequence}.  Your algorithm should have
  work $O(n)$ and span $O(\log n)$ where $n$ is the length of the
  array.  Do \emph{not} employ a sequential cut-off: your base case
  should process an array range containing one element.  Hint: Use
  this definition:
\begin{verbatim}
  class Result {
    int numLeftEdge;
    int numRightEdge;
    int numLongest;
    boolean entireRange;
    Result(int l, int r, int m, boolean a) {
      numLeftEdge=l; numRightEdge=r; numLongest=m; entireRange=a;
    }
  }
\end{verbatim}
For example, {\tt numLeftEdge} should represent the length of the
sequence at the \emph{beginning} of the range processed by a
subproblem.  Think carefully about how to combine results.
\item In English, describe how you would make your answer to part (a)
  more efficient by using a sequential cut-off.  In pseudocode, show
  the code you would use below this cut-off.
\end{enumerate}

\end{problem}

\newpage

\begin{problem}{\bf Fork-Join Parallelism: Leftmost Occurrence of Substring}

Consider the problem of finding the leftmost occurrence of the sequence
of characters {\tt cseRox} in an array of characters, returning the index
of the leftmost occurrence or {\tt -1} if there is none.  For example,
the answer for the sequence {\tt cseRhellocseRoxmomcseRox} is {\tt 9}.

\begin{enumerate}
\item In English (though some high-level pseudocode will probably help),
  describe a fork-join algorithm similar in design to your solution in problem
  1.  Use a sequential cut-off of at least 6 (the
  length of {\tt cseRox}) and explain why this
  significantly simplifies your solution.  Notice you still must deal
  with the leftmost occurrence being ``split'' across two recursive
  subproblems.
\item Give a much simpler fork-join solution to the problem that avoids the
  possibility of a ``split'' by using slightly overlapping
  subproblems.  Assume a larger sequential cut-off, for example 100.
  Give your solution precisely in pseudocode. Avoid off-by-one errors.
\end{enumerate}

\end{problem}

\newpage

\begin{problem}{\bf Amdahl's Law: Graphing the Pain}

  Use a graphing program such as a spreadsheet to plot the following
  implications of Amdahl's Law.  Turn in the graphs and tables with
  the data.

\begin{enumerate}
\item Consider the speed-up ($T_1/T_P$) where $P=256$ of a program
  with sequential portion $S$ where the portion $1-S$ enjoys perfect
  linear speed-up.  Plot the speed-up as $S$ ranges from .01 (1\%
  sequential) to .25 (25\% sequential).
\item Consider again the speed-up of a program with sequential portion $S$
  where the portion $1-S$ enjoys perfect linear speed-up.  This time,
  hold $S$ constant and vary the number of processors $P$ from 2 to
  32.  On the same graph, show three curves, one each for $S=.01$,
  $S=.1$, and $S=.25$.
\end{enumerate}

\end{problem}

\newpage

\begin{problem}{\bf Parallel Prefix and Pack}

  In this problem, the input is an array of strings and the output is
  an array of integers.  The output has the length of each string in
  the input, but empty strings are filtered out.  For example:
\begin{verbatim}
   [ "", "", "cse", "rox", "", "homework", "", "7", "" ]
\end{verbatim}
produces output:
\begin{verbatim}
   [ 3, 3, 8, 1]
\end{verbatim}
  A parallel algorithm can solve this problem in $O(\log n)$ span and
  $O(n)$ work by doing a parallel map to produce a bit vector, followed
  by a parallel prefix over the bit vector, followed by a parallel map
  to produce the output.

  Show the intermediate steps for the algorithm described above on the
  example above.  For each step, show the tree of recursive task
  objects that would be created (where a node's child is for two
  problems of half the size) and the fields each node needs.  Do not
  use a sequential cut-off.  Show three separate trees (for the three
  steps).  Explain briefly what each field represents.

  Note that because the input length is not a power of two, the tree
  will not have all its leaves at exactly the same height.
\end{problem}

\newpage

\begin{problem}{\bf Parallel Quicksort}

  Lecture presented a parallel version of quicksort with
  \emph{best-case} $O(\log^2 n)$ span and $O(n\log n)$ work.  This
  algorithm used parallelism for the two recursive sorting calls and
  the partition.

  \begin{enumerate}
  \item For the algorithm from lecture, what is the asymptotic \emph{worst-case}
  span and work.  Justify your answer.

  \item Suppose we use the parallel partition part of the algorithm,
  but perform the two recursive calls in sequence rather than in
  parallel.
  \begin{enumerate}
  \item[i.] What is the asymptotic best-case span and work?  Justify your answer.
  \item[ii.] What is the asymptotic worst-case span and work? Justify your answer.
  \end{enumerate}
  \end{enumerate}

\end{problem}

\newpage

\begin{problem}{\bf Another Wrong Bank Account}

Note: The purpose of this problem is to show you something you should
not do because it does not work.

Consider this pseudocode for a bank account supporting concurrent
access:
\begin{verbatim}
class BankAccount {
  private int balance = 0;
  private Lock lk = new Lock();
  int getBalance() {
    lk.acquire();
    int ans = balance;
    lk.release();
    return ans;
  }
  void setBalance(int x) {
    lk.acquire();
    balance = x;
    lk.release(); 
  }
  void withdraw(int amount) {
    lk.acquire(); 
    int b = getBalance();
    if(amount > b) {
      lk.release();
      throw new WithdrawTooLargeException();
    }
    setBalance(b - amount);
    lk.release(); 
  }
}
\end{verbatim}
The code above is wrong if locks are \emph{not} re-entrant.
Consider the \emph{absolutely horrible idea} of ``fixing'' this
problem by rewriting the {\tt withdraw} method to be:
\begin{verbatim}
  void withdraw(int amount) {
    lk.acquire(); 
    lk.release();
    int b = getBalance();
    lk.acquire();
    if(amount > b) {
      lk.release();
      throw new WithdrawTooLargeException();
    }
    lk.release();
    setBalance(b - amount);
    lk.acquire();
    lk.release(); 
  }
\end{verbatim}

\begin{enumerate}
\item Explain how this approach prevents blocking forever unlike the
  original code.
\item Show this approach is incorrect by giving an interleaving of two
  threads in which a withdrawal is forgotten.
\end{enumerate}

\end{problem}

\newpage

\begin{problem}{\bf Concurrent Queue with Two Stacks}

Consider this Java implementation of a queue with two stacks.  We do
not show the entire stack implementation, but assume it is correct.
Notice the stack has synchronized methods but the queue does not.
The queue is incorrect in a concurrent setting.
\begin{verbatim}
class Stack<E> {                              class Queue<E> {
  ...                                           Stack<E> in  = new Stack<E>();
  synchronized boolean isEmpty() { ... }        Stack<E> out = new Stack<E>();
  synchronized E pop() { ... }                  void enqueue(E x){ in.push(x); }
  synchronized void push(E x) { ... }           E dequeue() {
}                                                 if(out.isEmpty()) {
                                                    while(!in.isEmpty()) {
                                                      out.push(in.pop());
                                                    }
                                                  }
                                                  return out.pop();
                                                }
                                              }
\end{verbatim}
\begin{enumerate}
\item Show the queue is incorrect by showing an interleaving that
  meets the following criteria:
  \begin{enumerate}
  \item[i.] Only one thread ever performs {\tt enqueue} operations and
  that thread enqueues numbers in increasing order (1, 2, 3, ...).
  \item[ii.] There is a thread that performs two {\tt dequeue} operations
  such that its first {\tt dequeue} returns a number larger than its
  second {\tt dequeue}, which should never happen.
  \item[iii.] Every {\tt dequeue} succeeds (the queue is never empty).
  \end{enumerate}
  Your solution can use 1 or more additional threads that perform {\tt
    dequeue} operations.
\item A simple fix would make {\tt enqueue} and {\tt dequeue}
  synchronized methods.  Explain why this would never allow an {\tt
  enqueue} and {\tt dequeue} to happen at the same time.
\item To allow an {\tt enqueue} and a {\tt dequeue} to
  operate on a queue at the same time (at least when {\tt out} is not empty),
  we could try either of the approaches below for {\tt
    dequeue}.  For each, show an interleaving that demonstrates the
  approach is broken.  Your interleaving should satisfy the three
  properties listed in part (a).
\begin{verbatim}
  E dequeue() {                                E dequeue() {
    synchronized(out) {                          synchronized(in) {
      if(out.isEmpty()) {                          if(out.isEmpty()) {
        while(!in.isEmpty()) {                       while(!in.isEmpty()) {
          out.push(in.pop());                          out.push(in.pop());
        }                                            }
      }                                            }
      return out.pop();                           }
    }                                             return out.pop();
  }                                            } 
\end{verbatim}
\item Provide a solution, based on two stacks as above, that correctly allows
an {\tt enqueue} and a {\tt dequeue} to happen at the same time, at
least when {\tt out} is not empty.  Your solution should define {\tt dequeue} and
involve multiple locks.
\end{enumerate}

\end{problem}

\newpage

\begin{problem}{\bf Simple Concurrency with B-Trees}

  Note: Real databases and file systems use very fancy fine-grained
  synchronization for B-Trees such as ``hand-over-hand locking''
  (which we did not discuss), but this problem considers some
  relatively simpler approaches.

  Suppose we have a B Tree supporting operations {\tt insert} and {\tt
    lookup}.  A simple way to synchronize threads accessing the tree
  would be to have one lock for the entire tree that both operations
  acquire/release.
\begin{enumerate}
\item Suppose instead we have one lock per node in the tree. Each
  operation acquires the locks along the path to the leaf it needs
  and then at the end releases all these locks.  Explain why this
  allegedly more fine-grained approach provides absolutely no benefit.
\item Now suppose we have one readers/writer lock per node and {\tt
    lookup} acquires a read lock for each node it encounters whereas
  {\tt insert} acquires a write lock for each node it encounters.  How
  does this provide more concurrent access than the approach in part
  (a)?  Is it any better than having one readers/writer lock for the
  whole tree (explain)?
\item Now suppose we modify the approach in part (b) so that {\tt
    insert} acquires a write lock \emph{only for the leaf node} (and
    read locks for other nodes it encounters).  How would
  this approach increase concurrent access?  When would this be
  \emph{incorrect}?  Explain how to fix this approach without changing the
  asymptotic complexity of {\tt insert} by detecting when it is
  incorrect and in (only) those cases, starting the {\tt insert} over
  using the approach in part (b) for that {\tt insert}.  Why would
  reverting to the approach in part (b) be fairly rare?
\end{enumerate}

\end{problem}

\newpage

\begin{problem}{\bf Concurrent/Parallel Graph Traversal}

\begin{enumerate}
\item In Java or pseudocode, define an \emph{unbounded stack} with
  operations {\tt push} and {\tt pop} that can be used by threads
  concurrently.  Make the element type generic and use a linked list
  for the underlying implementation.  A {\tt pop} should not raise an
  error.  Instead it should wait until the stack is not empty.  
  Use a condition variable.  (Java detail: The {\tt wait}
  method throws an exception that Java requires you catch, but it
  doesn't matter if you include this detail in your solution.)
\item The method {\tt traverseFrom} in the code below uses your answer
  to part (a) to apply the {\tt doIt} method to every node reachable
  from some node in a graph.  The order {\tt doIt} is applied to nodes
  does not matter.  The code uses 4 threads to try to improve
  performance.  All you need to do is write the {\tt run} method
  (probably around 10 lines) for {\tt GraphWalker} using the other
  code.  Requirements:
  \begin{itemize}
  \item No node should have {\tt doIt} applied to it more than once.
  You can assume the {\tt visited} field of each node is initially
  {\tt false} for this purpose.
  \item Use the stack to hold nodes that still need to be processed.
  That way threads not busy completing a {\tt doIt} call can find
  useful work.
  \item Avoid all synchronization errors.  (Hint: Your code does not
  need any \emph{explicit} synchronization since it can call other
  code that already performs synchronization.)  
  \item Because {\tt doIt} may be expensive, do not hold any locks while
  calling it.
  \end{itemize}
  Note that after all reachable nodes have been processed, all threads
  will be blocked forever waiting for the stack to become non-empty.
  That is fine for a homework problem.  Also note that you
  may or may not find {\tt isVisited} helpful.
\end{enumerate}

\begin{tabular}{ll}
\begin{minipage}[t]{.48\linewidth}
\begin{verbatim}
class GraphNode {
  boolean visited = false;
  GraphNode[] neighbors; 
  // ... other fields, constructor, 
  // etc. omitted ...
  synchronized boolean isVisited() { 
    return visited; 
  }
  synchronized boolean wasAndSetVisited(){ 
    boolean ans = visited;
    visited = true;
    return ans;
  }
}
\end{verbatim}
\end{minipage}
&
\begin{minipage}[t]{.48\linewidth}
\begin{verbatim}
class GraphWalker extends java.lang.Thread {
  Stack<GraphNode> stk;
  GraphWalker(Stack<GraphNode> s) {
    stk = s;
  }
  void doIt(GraphNode n) {/*...*/}

  // ... FOR YOU ...

}
class Main {
  public static final int NUM_THREADS = 4;
  void traverseFrom(GraphNode root) {
    Stack<GraphNode> stk = 
       new Stack<GraphNode>();
    stk.push(root);
    for(int i = 0; i < NUM_THREADS; i++)
      new GraphWalker(stk).start();
  }
}
\end{verbatim}
\end{minipage}
\end{tabular}

\end{problem}

\end{document}