### 2023-06-08 CAIE/shims.tex

% !TeX spellcheck = en_US
%\documentclass[preprint,authoryear]{elsarticle}
\documentclass[preprint]{elsarticle}
% sudo apt-get install -y texlive-publishers

\usepackage[left]{lineno}
\usepackage[latin1]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{multirow}

\usepackage{lineno,hyperref}
\modulolinenumbers[5]

\usepackage[none]{hyphenat}

\usepackage{url}
\usepackage{booktabs}

\usepackage{tikz}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}

\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{amsthm}

\usepackage{caption}
\captionsetup{labelfont=bf}
\captionsetup{skip=0pt}

\usepackage{pgfplots}
\pgfplotsset{compat=newest}
\usepackage{algorithm}

\usepackage[noend]{algpseudocode}

\usepackage[margin=0.9in]{geometry}

\usepackage[normalem]{ulem}
%\usepackage{xcolor}

\newcommand{\Comentar}[1]{\State {\cmmt{#1}}}
\newcommand{\Break}{\State \textbf{break}}
\renewcommand{\Return}{\State \textbf{return}~}
\renewcommand{\algorithmicensure}{\textbf{Parameters:}}

\renewcommand\algorithmicthen{}
\renewcommand\algorithmicdo{}

\algblock{ForEach}{EndFor}
\algblock{ForDownTo}{EndFor}
\algblock{ForTo}{EndFor}

\newcommand{\Block}[1]{\State #1 \{}
\newcommand{\EndBlock}{\State \}}

\usepackage{scrextend}
\newcommand{\boldm}[1] {\mathversion{bold}#1\mathversion{normal}}
\newcommand{\round}[1]{\ensuremath{\lfloor#1\rceil}}
\usepackage{setspace}
\usepackage{array}
\usepackage{color, colortbl}
\definecolor{Gray}{gray}{0.9}
\hyphenation{cost-ef-fec-tive-ness}

\newcommand{\specialcell}[2][c]{%
	\begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}

\usepackage{textcomp}

\journal{Computers \& Industrial Engineering}

\begin{document}

\begin{algorithm}[H]
	\caption{$Shims$\/ heuristic at the node $\pi_k$}  \label{alg:shims}
	\begin{algorithmic}[1]

		\Procedure{$SolveNode$}{$Shims, \pi_k, G, time$}

		\State $T_{begin} \gets$ current system time

		\State Let $G(M \cup N_{\pi_k} \cup Q_{\pi_k}, E_{Q_{\pi_k}})$   \label{shims:initQ}

		\State Sort $M$ by $|D_i^{long}|$ in non-descending order \label{shims:pallets}

		\State $E_{N_{\pi_k}} \gets \varnothing$  \Comment{An empty set of pallet-item edges}

		\State $\tau_{\pi_k} \gets 0$	 \Comment{Torque at the node $\pi_k$}	
		\State $\tau_{max} \gets W_{max} \times limit^{CG}_{long}$ \Comment{Maximum aircraft torque}

		\State $vol_i \gets 0$, $1\leq i \leq m$

		\State $limit \gets 0.92$  \label{shims:limit}

		\For{$i \gets 1$ to $m$}  \label{shims:eta1_a}

			\For{$q \gets 1$ to $m_{\pi_k}$}

				\If{$(i, q) \in E_{Q_{\pi_k}}$} 
					\State $vol_i \gets vol_i + v_q$ 
					\State $\tau_{\pi_k} \gets \tau_{\pi_k} + w_q \times D_i^{long}$
				\EndIf	
			\EndFor	
			\State Let $E$\/ be an array of $n_{\pi_k}$ possibles edges of the pallet $i$\/ sorted by $\theta_{ij}^{\pi_k}$ in non-ascending order \label{shims:possible}
			\State $\eta_1 \gets 1$  
			\Repeat
				\State $e_{ij} \gets E_{\eta_1}$ \Comment{$e_{ij}$\/ is the allocation of the item $j$\/ to the pallet $i$}

				\If{ ($E_{N_{\pi_k}} \cup \{e_{ij}\}$ is feasible) {\bf and} ($vol_i \leq V_i \times limit$) {\bf and} ($| \tau_{\pi_k} + w_j \times D_i^{long} | \leq W_{max} \times limit^{CG}_{long}$)}

					\State $E_{N_{\pi_k}} \gets E_{N_{\pi_k}} \cup \{e_{ij}\}$

					\State $vol_i \gets vol_i + v_j$

					\State $\tau_{\pi_k} \gets \tau_{\pi_k} + w_j \times D_i^{long}$

					\State $\eta_1 \gets \eta_1 + 1$

				\EndIf
			\Until ($vol_i > V_i \times limit$) {\bf or} ($\eta_1>n_{\pi_k}$)		 \Comment{End of the first phase} \label{shims:endfirst}

			\State $slack_i \gets V_i - vol_i$  \label{shims:beginsecond}
			\State $\eta_2 \gets \eta_1$  \label{shims:eta2a}
			\While{($\eta_2 \leq n_{\pi_k}$) {\bf and} ($ vol_i < (1 + 3 \times (1-limit)) \times V_i)$)}

					\State $e_{ij} \gets E_{\eta_2}$

					\State $vol_i \gets vol_i + v_j$
					\State $\eta_2 \gets \eta_2 + 1$  \label{shims:eta2} \label{shims:endsecond} \Comment{End of the second phase}
			\EndWhile

			\State $vol \gets 0$, $b \gets 1$, $shims_b \gets \varnothing$, {\color{blue} $Set \gets \{ shims_b \}$ } \label{shims:beginthird}

			\For{$x \gets \eta_1$ to $\eta_2$} \label{edges_indexes}

				\If{ $T_{current} - T_{begin} > time$}
				\State {\bf break}  \Comment{Time limit exceeded}
				\EndIf

				\State $NewShims \gets$ {\bf True} \label{new_shims}
				\State $e_{ij} \gets E_x$

				\For{$shims \in Set$} \label{shims_set}

					\If {($e_{ij} \not\in (E_{N_{\pi_k}} \cup shims))$ {\bf and} ($e_{ij}$ is feasible)  {\bf and} ($(v_j + vol) \leq slack_i$)}

						\State $shims \gets shims \cup \{e_{ij}\}$
						\State $vol \gets vol + v_j$
						\State $NewShims \gets$ {\bf False} \label{new_shims_false}

						\State {\bf break}  
					\EndIf

				\EndFor

				\If{$NewShims$} \label{new_shims2}
					\State $vol \gets 0$
					\State $b \gets b + 1$
					\State $shims_b \gets \{\ \}$
					\State $shims_b \gets shims_b \cup \{e_{ij}\}$
					\State $Set \gets Set \cup \{\ shims_b\ \}$
				\EndIf

			\EndFor

			\State $sh_w \gets shims$, where $shims \in Set$ and $\sum_{e_{ab} \in shims} w_b$ is maximum  \label{best_weight}		
			\State $sh_v \gets shims$, where $shims \in Set$ and $\sum_{e_{ab} \in shims} v_b$ is maximum \label{best_volume}
			\State $sh_{best} \gets shims$, where $shims \in \{sh_w, sh_v\}$ and $\sum_{e_{ab} \in x} s_b$ is maximum \label{best_score}

			\State $E_{N_{\pi_k}} \gets E_{N_{\pi_k}} \cup sh_{best}$ 	\label{shims:endthird}   \Comment{End of the third phase}

		\EndFor

		\Return $G(M \cup N_{\pi_k} \cup Q_{\pi_k}, E_{Q_{\pi_k}} \cup E_{N_{\pi_k}})$

		\EndProcedure
	\end{algorithmic}
\end{algorithm}

Initially, $Q_{\pi_k}$\/ (line \ref{shims:initQ}) corresponds to the packed contents that remain on board. It is important to remember that $E_{Q_{\pi_k}}$\/ and $M$\/ were modified by $UpdatePacked(M, Q_{\pi_k}, \pi_k)$ and $SetPalletsDestinations(M, \pi_k)$. Then, the pallets $i$\/ are considered in non-descending order of $|D_i^{long}|$.

For each pallet $i$, its $n_{\pi_k}$\/ possible edges $e_{ij}^{\pi_k}$\/ are considered in non-increasing order of $\theta_{ij}^{\pi_k}$:

{\color{blue}

\begin{itemize}

	\item In the {\it greedy phase}\/ (lines \ref{shims:pallets}-\ref{shims:endfirst}), a partial solution for each pallet $i$\/ is constructed by edges inclusion, until its volume $V_i \times limit$\/ is reached. The value $limit=0.92$\/ (line \ref{shims:limit}) was empirically determined by the {\it iRace} tool \cite{LopezIbanezManuel2016}.

	From lines \ref{shims:limit} to \ref{shims:endfirst}, as items are inserted, the value of $\eta_1$ is incremented and stops when the pallet volume reaches 92\% of its capacity.

	When it stops, this is the initial value of $\eta_2$, line \ref{shims:eta2a}.
	$\eta_2$, then it is incremented until the volume exceeds the pallet capacity with a factor of 1 + ( 2 x [1 - limit]).
	This factor of was obtained empirically, through many test with different problem instances.
	Higher values can improve the solution, but at the expense of performance. So we tried to keep it as low as possible, a trade-off between quality and performance.

	\item In the {\it composition phase}\/ (lines \ref{shims:beginsecond}-\ref{shims:endsecond}), a set of shims named $Set$\/ is created for each pallet $i$, where each shim is formed by a set of edges in the range $[\eta_1,\eta_2]$, whose total volume is limited by $slack_i$.

	After the $\eta_2$ index, there may be items with little probability of being in the optimal solution because their low value of $\theta^{\pi_k}_{ij}$, being not considered for inclusion.

	In line \ref{shims:beginthird} the initial $Shim$ and $Set$ are created.

	In this phase, the heuristic that provided the best results, both in terms of time and quality, is based on {\it First-Fit Decreasing}, which is an approximation algorithm for the {\it Bin Packing Problem}\/ \cite{JohnsonGarey1985}. Basically, shims are created by accumulating the following edges, taking $slack_i$\/ as a limit.

	\item In the {\it selection phase}\/ (lines \ref{shims:beginthird}-\ref{shims:endthird}), the best shim in $Set$\/ is chosen. Initially, two shims are found: $sh_w$\/ with larger weight, and $sh_v$\/ with larger volume. Between the two, the one with the highest score will be chosen and its edges are inserted into $E_{N_{\pi_k}}$.

\end{itemize}

}

\end{document}